\subsection{Monocular Hand Pose and Shape Estimation}\label{subsec:sota_estimation}

\subsubsection{Model-Free Keypoint Estimation}\label{subsubsec:model-free}

Model-free methods approach monocular hand tracking by training a network to
regress joint coordinates directly from the input image, without fitting
a parametric model at inference time.
Their principal attraction is computational efficiency: the output space is
compact and inference is fast, enabling real-time performance on commodity
mobile devices using a standard RGB camera and requiring no specialised
depth-sensing hardware~\cite{zhang2020mediapipehands}.

Spurr et al.~\cite{spurr2020weakly} proposed a weakly supervised approach to
3D hand pose estimation that incorporates biomechanical constraints---including
bone length constraints, a palm-structure term penalising invalid curvature and
inter-bone angles among the metacarpal root bones,
and anatomical joint angle limits encoding the valid range
of flexion and abduction---directly into the training objective as
differentiable soft losses.
Rather than fitting a kinematic or parametric model to the predictions, the
method extracts the relevant biomechanical quantities directly from the
predicted 3D joint positions in order to constrain them during training.
This allows the network to leverage images annotated only with 2D keypoints as
a supplementary supervisory signal, reducing reliance on per-frame 3D
ground-truth annotations and improving generalisation to real images.
The method additionally estimates the absolute depth of the root joint,
$Z_\mathrm{root}$, from the underlying 2.5D representation, further stabilised
by a dedicated refinement network.
As with MediaPipe, the output is a set of 3D joint coordinates
$\mathbf{J}^{3D} \in \mathbb{R}^{21\times3}$: no parametric shape model is
fitted and no mesh is produced.

The common limitation of both approaches is that their output consists solely
of joint positions in camera space.
As detailed in~\autoref{subsec:mediapipe}, the depth coordinate is wrist-relative
and carries no absolute metric scale.
Neither method encodes joint rotations or a kinematic hierarchy in its output:
both produce a flat set of 3D joint positions, a limitation introduced
in~\autoref{subsec:gap}.
This coordinate output suffices for gesture classification---as the finger-state
mechanism described in~\autoref{subsec:mediapipe} illustrates---but it cannot
directly animate a rigged skeletal avatar or generate motion logs in a
canonical, scale-invariant representation.



\subsubsection{Model-Based Regression}\label{subsubsec:model-based}

A second family of methods produces anatomically grounded output by grounding
network predictions in the MANO parametric hand model.
Given pose coefficients $\boldsymbol{\theta}$ and shape coefficients
$\boldsymbol{\beta}$, the full 3D mesh follows analytically from the MANO forward
pass, yielding output that is anatomically valid by construction, invariant to
hand size, and directly compatible with graphics and motion-analysis pipelines.
Methods in this family differ in how they arrive at these parameters: some regress
them directly from image features, while others derive them indirectly via
predicted 3D joint positions.

Zhou et al.\ \cite{zhou2020monocular} presented a method for real-time monocular
hand shape and motion capture that runs at over 100\,fps, following the indirect
route.
The system comprises two learned modules.
A joint detection network, DetNet, built on a ResNet-50 backbone, regresses
root-relative, scale-normalised 3D joint positions from the input image under a
multi-task scheme that simultaneously exploits 2D-annotated real images and
3D-annotated synthetic data.
A subsequent inverse kinematics network, IKNet, then solves the inverse kinematics
problem in a single feed-forward pass through a seven-layer fully-connected
network trained on motion-capture data, converting the predicted joint positions
into per-joint rotations expressed as unit quaternions.
These rotations drive the MANO model directly, while shape parameters
$\boldsymbol{\beta}$ are estimated separately by minimising the discrepancy
between MANO bone lengths and the bone lengths derived from the 3D joint
predictions.
On an NVIDIA GTX\,1080\,Ti, DetNet requires 8.9\,ms and IKNet 0.9\,ms per
frame, yielding a combined throughput exceeding 100\,fps.

Park et al.\ \cite{park2022handoccnet} presented HandOccNet, addressing the
robustness challenge of partial occlusion within the direct model-based regression
framework.
The backbone separates image features into primary features and secondary features
using a learned necessity map, where primary features correspond to
high-importance, hand-relevant regions and secondary features correspond to
low-importance regions associated with occlusion.
Two successive Transformer-based modules then operate on these features: the
Feature Injecting Transformer (FIT) injects information from the primary features
into the locations of the secondary features by treating the secondary features as
queries and the primary features as key--value pairs, while the Self-Enhancing
Transformer (SET) subsequently refines the result via self-attention.
The enriched feature map is passed to a regressor that predicts MANO pose
parameters $\boldsymbol{\theta} \in \mathbb{R}^{48}$ and shape parameters
$\boldsymbol{\beta} \in \mathbb{R}^{10}$ directly, from which the final 3D hand
mesh is obtained via the MANO layer.
HandOccNet was shown to outperform all prior works on the HO-3D and FPHA
hand-object interaction benchmarks.

Pavlakos et al.\ \cite{pavlakos2024hamer} presented HaMeR (Hand Mesh Recovery),
a fully transformer-based approach using a large-scale Vision Transformer (ViT-H)
backbone followed by a transformer decoder head that regresses MANO pose and
shape parameters directly from a monocular RGB image.
The model is trained on 2.7\,M annotated examples consolidated from ten
datasets providing either 2D or 3D hand annotations; the authors note that the
majority of these examples are collected in controlled studio or multi-camera
environments, with only 5\% drawn from in-the-wild sources.
HaMeR achieves state-of-the-art accuracy across most metrics on the FreiHAND and
HO-3D benchmarks and, when evaluated on the authors' in-the-wild HInt benchmark,
yields 2--3$\times$ improvements in PCK@0.05 over prior methods, with the ablation
study demonstrating that both the increased data scale and the high-capacity
architecture independently contribute to this gain.

The shared representational commitment across all three systems is the use of MANO
as the output representation, which enforces anatomical validity and enables direct
compatibility with downstream graphics pipelines.
The systems differ, however, in how MANO parameters are obtained: Zhou et al.\
follow an indirect route in which the backbone first predicts 3D joint positions
that are subsequently converted to rotations and shape parameters via IKNet and a
bone-length optimisation; HandOccNet and HaMeR instead regress MANO parameters
directly from image features in a single feed-forward pass.
Zhou et al.\ explicitly report their system achieving over 100\,fps on a dedicated
GPU; the papers of Park et al.\ and Pavlakos et al.\ do not report inference
timing or hardware requirements.



\subsubsection{Real-Time Tracking on Accessible Hardware}\label{subsubsec:sota_hardware}

Efforts to reduce the computational footprint of hand mesh reconstruction have
produced systems that do achieve interactive rates on mobile hardware.
Chen et al.\ \cite{chen2022mobrecon} presented MobRecon, which runs at 83\,fps
on an Apple A14 CPU and achieves substantially lower inference cost than prior
graph-based or transformer-based regressors.
However, MobRecon is a non-parametric, vertex-based method: rather than
regressing \acrshort{mano} pose and shape parameters, it directly predicts the
3D coordinates of the 778 mesh vertices.
This design choice is what enables its efficiency---the authors explicitly note
that \acrshort{mano} coefficient estimation is ill-suited to lightweight networks
because it is a highly abstract problem that ignores spatial
correlations~\cite{chen2022mobrecon}---but it means that MobRecon does not
produce joint rotations in a kinematic hierarchy and therefore cannot directly
drive a rigged skeletal avatar without a separate pose-extraction step.

MediaPipe Hands~\cite{zhang2020mediapipehands} achieves real-time landmark
tracking on commodity CPU hardware but, as established
in~\autoref{subsubsec:model-free}, its raw coordinate output lacks
the canonical kinematic hierarchy.

Both MobRecon and MediaPipe thus reach interactive rate on accessible hardware
only by forgoing the \acrshort{mano} joint rotations needed to drive a rigged
skeleton.
Closing that gap without a learned pose model is the route the present system
takes; the solver families it draws on are surveyed in~\autoref{subsec:sota_ik},
and its position relative to the literature is set out
in~\autoref{subsec:sota_positioning}.
